A New Model to Predict Survival in Patients with End-Stage Liver Disease

### 2022-07-18 analysis/article.Rmd
refactor: replace gaertner2022 by sealock2022

Liver cirrhosis is the terminal result of fibrotic remodeling
of liver tissue due to chronic damage.
This end-stage of liver disease is usually irreversible and
the only available therapy is liver transplantation.
However, the shortage in grafts for transplantation from deceased donors
requires risk stratification and precise and fair allocation rules.
The allocation of liver transplantation in most countries is based on
disease severity determined by the model of end-stage liver disease (MELD)
[@malinchoc2000; @wiesner2003; @optn_policies2021].
The MELD score estimates the patients' 3-month mortality risk based on
laboratory results, namely bilirubin, creatinine and international normalized
ratio (INR).
The MELD score has been extended by the sodium level
(MELD-Na score) as this was found to be an important additional risk factor in
liver cirrhosis [@kim2008; @optn_policies2021].
The MELD was initially developed for predicting the survival of patients
undergoing transjugular intrahepatic portosystemic shunts.
Afterwards it was revalidated for predicting mortality risk in patients awaiting
a liver transplantation.
Although the MELD score should be an objective allocation score especially the
creatinine and INR are highly dependent on the utilized laboratory methods
[@trotter2004; @cholongitas2007]. Patients with identical disease state could
have very different MELD scores and thus get different priority on the liver
transplantation waiting list.
Furthermore often, e.g. for acute-on-chronic liver failure, the MELD score
underestimates the mortality risk [@hernaez2020].

@@

It has shown that women receive in median lower scores in MELD-Na and therefore
are less likely getting an urgent listing for liver transplantation
[@gaertner2022].
To address sexual differences in laboratory measurements we applied the *zlog*
transformation [@hoffmann2017] in our study.
Additionally this transformation yield comparable laboratory measurements
across different clinical laboratories and methods.
It also reduce the influence of age- and sex-related differences,
which could improve the objectivity of the allocation process.
The already nearly normal scaled *zlog*-transformed laboratory measurements
allow us to avoid the sample depending scaling prior to model development.
Both should increase the external applicability of our approach.

>>>>

Liver cirrhosis is the terminal result of fibrotic remodeling
of liver tissue due to chronic damage.
This end-stage of liver disease is usually irreversible and
the only available therapy is liver transplantation.
However, the shortage in grafts for transplantation from deceased donors
requires risk stratification and precise and fair allocation rules.
The allocation of liver transplantation in most countries is based on
disease severity determined by the model of end-stage liver disease (MELD)
[@malinchoc2000; @wiesner2003; @optn_policies2021].
The MELD score estimates the patients' 3-month mortality risk based on
laboratory results, namely bilirubin, creatinine and international normalized
ratio (INR).
The MELD score has been extended by the sodium level
(MELD-Na score) as this was found to be an important additional risk factor in
liver cirrhosis [@kim2008; @optn_policies2021].
The MELD was initially developed for predicting the survival of patients
undergoing transjugular intrahepatic portosystemic shunts.
Afterwards it was revalidated for predicting mortality risk in patients awaiting
a liver transplantation.
Although the MELD score should be an objective allocation score especially the
creatinine and INR are highly dependent on the utilized laboratory methods
[@trotter2004; @cholongitas2007].
Additionally women are disadvantage [@sealock2022].
Patients with identical disease state could have very different MELD scores and
thus get different priority on the liver transplantation waiting list.
Furthermore often, e.g. for acute-on-chronic liver failure, the MELD score
underestimates the mortality risk [@hernaez2020].

@@

It has shown that women receive in median lower scores in MELD-Na and therefore
are less likely getting an urgent listing for liver transplantation
[@sealock2022].
To address sexual differences in laboratory measurements we applied the *zlog*
transformation [@hoffmann2017] in our study.
Additionally this transformation yield comparable laboratory measurements
across different clinical laboratories and methods.
It also reduce the influence of age- and sex-related differences,
which could improve the objectivity of the allocation process.
The already nearly normal scaled *zlog*-transformed laboratory measurements
allow us to avoid the sample depending scaling prior to model development.
Both should increase the external applicability of our approach.
